Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge

### 2025-02-08 paper/main.tex
Updates

\subsection{Ranking Potential New Edges}

>>>>

In this section, we give some of the practical aspects of the
\emph{ExpertInLoop} algorithm, and then compare the algorithm with automated
causal discovery algorithms.

\subsection{Ranking Potential New Edges}
\begin{algorithm}[h]
\DontPrintSemicolon
\SetAlgoLined
\SetKwFunction{RankedExpand}{RankedExpand}
\SetKw{KwGoTo}{go to}
\SetKwProg{Fn}{Function}{:}{}
\Fn{{\sc RankedExpand}($V,E,\mathcal{D}_G,\mathcal{A}_G$,$\phi$,$B=\emptyset$)}{
    let $\phi_{\max} \gets 0 $\;
    \ForEach{$X, Y$ where $X \to Y \notin E$ and $Y \to X \notin E$ and 
	$\{X,Y\} \notin B$}{
        let $\mathbf{Z}$ be a set that $d$-separates $X$ and $Y$ in $(V,E)$\;
        \If{$\mathcal{D}_G(X, Y, \mathbf{Z}) = 0$}{
		\If{$\lvert \phi(X, Y, \mathbf{Z}) \rvert > \phi_{\max}$}{
			$ \phi_{\max} \gets \lvert \phi(X, Y, \mathbf{Z}) \rvert $ \;
			$ L_{\max} \gets \mathcal{A}_G(X, Y)$ \;
		}
        }
    }
    add $L_{\max}$ to $E$\;
    \Return{$(V,E)$}\;
}
\caption{Adding an edge between variables with the highest correlation}
\label{algo:ranked_expand}
\end{algorithm}

In Algorithm~\ref{algo:expert}, in each iteration we add a new edge to the DAG
chosen by Algorithm~\ref{algo:expand}. This edge is chosen at random depending
on the order of iteration in Algoirthm~\ref{algo:expand}. A more optimal way to
add these edges would be to choose the pairs of variables that could help
improve the fit the most in each step. One potential greedy approach to do this
would be to select the variable pair $ X $ and $ Y $ that have the highest
amount of observed conditional correlation in the data. In
Algorithm~\ref{algo:ranked_expand}, we slightly modify
Algorithm~\ref{algo:expand} to show this approach. In this case we first go
through all potential edges that can be added to the DAG in each step, and rank
them based on the conditional association between them given the current DAG.
This gives us a measure of how much of the observed correlation between them is
not explained by the current model. We then choose the pair of variables the
highest conditional association. This modification allows us to prioritize pair
of variables to quickly improve the fit of the model. For the conditional
measure of association, we use a measures outlined in
Section~\ref{sec:background} depending on the data type of the variables.
